---
title: "20231006-male-and-female-DiNV-injections-rep-2"
format: gfm
editor: visual
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width=7)
```

Load in packages needed for the analysis

```{r}
library("survival")
library("survminer")
```

### Loop to convert the example data.frame 'df' into properly formatted data.frame 'results'

```{r}
#write a function to transform a data.frame that has the column format 'vial | treatment | D0 | D1 | D2...', with one row for each vial
#into a long version in tidy format that can be input to make a survivorship curve
convert_df<-function(df){
  #open empty data.frame to store results
  results<-data.frame(vial=character(),treatment=character(),dead=numeric(),status=numeric())
  #This loop will index out each row (one row per vial) one at a time, transform it into long format (one row per fly), and add the information to the empty data.frame called results
  for(i in 1:nrow(df)){
  #isolate the row (vial) you want to work on
  temp<-df[i,]
    #iteratively isolate each day for this vial (day 0 must be column 3, day 1 column 4, etc.). Loop stops the column before the last day
    for(j in 3:(ncol(temp)-1)){
      #assign the number of flies that died in the vial on that day (starting with day 1) to the variable 'z'
      z<-temp[1,j]-temp[1,j+1]
        #if >0 flies died add this information to the results dataframe
        if(z>0){
          #iterate over 1 through total number of dead flies
          for(k in 1:z){
            #add a new row to the 'results' data.frame for the given dead fly, specifying vial #, treatment, day died, and
            #record the current vial #
            vial<-temp[,1]
            #record the genotype of the current vial
            treatment<-temp[,2]
            #record the death date of the flies that died on this day (assumes that your input DF starts with day 0 in column 3)
            dd<-j-2
            #append this information into a new row in the 'results' data.frame, and add a '1' in the 4th column to indicate mortality
            results[nrow(results)+1,]<- c(vial,treatment,dd,1)
          } #close for loop
        } #close if loop
    } #close for loop
  
  #now assign the number of flies remaining in the vial on the last day (value in the last column of the row) to the variable 'z'
  z<-temp[1,j+1]
    #if there are any flies alive in the vial on the last day
    if(z>0){
      #iterate over 1:(number of flies alive on the last day)
      for(l in 1:z){
        #record the current vial #
        vial<-temp[,1]
        #record the genotype of the current vial
        treatment<-temp[,2]
        #record the last day we recorded this fly alive (assumes that your input DF starts with day 0 in column 3)
        dd<-j-2
        #append this information into a new row in the 'results' data.frame, and add a '0' in the 4th column to indicate that the fly made it to the end of the experiment
        results[nrow(results)+1,]<- c(vial,treatment,dd,0)
      } #close for loop
    } #close if loop
  } #close original for loop
results$dead<-as.numeric(results$dead)  #reiterate that this column must be class numeric
results$status<-as.numeric(results$status)  #reiterate that this column must be class numeric
results$vial <- as.factor(results$vial) # make sure vial is considered a factor
# gives you only the results dataframe as output from function 
return(results) 
} #close function
```

Read in raw data

```{r}
#read the file from csv
df<-read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/Infection_survival_analyses/20231006-rep-2-male-female-nanoject-16Cq/20231006-counts.csv")

# no need to cut down on the file because I already separated out to to be the columns needed 
```

### Convert each of these dataframes to long and tidy format using function defined above

```{r}
df.convert<-convert_df(df)
```

Plot the survival curve

```{r}
# change to not have confidence intervals in this one so you can see them 
df_fit<- survfit(Surv(dead, status) ~ treatment, data=df.convert)
ggsurvplot(df_fit,
          pval = TRUE, conf.int = FALSE,
          #risk.table = TRUE, # Add risk table
          #risk.table.col = "strata", # Change risk table color by groups
          #linetype = "strata", # Change line type by groups
          #surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("orchid", "aquamarine", "blueviolet", "darkslategray3")) + ylab("Survival Proporation") + xlab("Days post injection")

```

```{r}
df_fit_t<- coxph(Surv(dead, status) ~ treatment, data=df.convert)
summary(df_fit_t)
```

### look at combined data from rep one and two of this infection experiment 

Read in data and convert
```{r}
#read the file from csv
df2<-read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/Infection_survival_analyses/20231006-rep-2-male-female-nanoject-16Cq/male-female-rep-1-2-counts-combo.csv")

# no need to cut down on the file because I already separated out to to be the columns needed 

df2.convert<-convert_df(df2)
# add in a column for block/replicate
# repeate the letter A 69 times (number of rows for block 1), then after that repeate B 69 times (number of rows for block 2)
df2.convert$Block <- rep(c("A","B"), c(69, 69))

```


Plot the survival curve without the blocks separated

```{r}
# change to not have confidence intervals in this one so you can see them 
df2_fit<- survfit(Surv(dead, status) ~ treatment, data=df2.convert)
ggsurvplot(df2_fit,
          pval = TRUE, conf.int = FALSE,
          #risk.table = TRUE, # Add risk table
          #risk.table.col = "strata", # Change risk table color by groups
          #linetype = "strata", # Change line type by groups
          #surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c("orchid", "aquamarine", "blueviolet", "darkslategray3")) + ylab("Survival Proporation") + xlab("Days post injection")

```


Plot the survival curve with the blocks separated

```{r}
# change to not have confidence intervals in this one so you can see them 
df3_fit<- survfit(Surv(dead, status) ~ treatment + Block, data=df2.convert)
ggsurvplot(df3_fit,
          pval = TRUE, conf.int = FALSE,
          #risk.table = TRUE, # Add risk table
          #risk.table.col = "strata", # Change risk table color by groups
          #linetype = "strata", # Change line type by groups
          #surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw() # Change ggplot2 theme
          ) + ylab("Survival Proporation") + xlab("Days post injection")
```

Model with block as a factor 

```{r}
df_fit4<- coxph(Surv(dead, status) ~ treatment + Block, data=df2.convert)
summary(df_fit4)
```

Not sure what this is comparing here, maybe I need to separate out sex differently?






