---
title: "20230509 innubila 16 Cq DiNv"
output: github_document
date: '2023-06-01'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width=9, fig.height = 7)
```

Load in packages needed for the analysis

```{r}
library("survival")
library("survminer")
library(tidyr)
library(stringr)
```


### Loop to convert the example data.frame 'df' into properly formatted data.frame 'results'
```{r}
#write a function to transform a data.frame that has the column format 'vial | treatment | D0 | D1 | D2...', with one row for each vial
#into a long version in tidy format that can be input to make a survivorship curve
convert_df<-function(df){
  #open empty data.frame to store results
  results<-data.frame(vial=character(),treatment=character(),dead=numeric(),status=numeric())
  #This loop will index out each row (one row per vial) one at a time, transform it into long format (one row per fly), and add the information to the empty data.frame called results
  for(i in 1:nrow(df)){
  #isolate the row (vial) you want to work on
  temp<-df[i,]
    #iteratively isolate each day for this vial (day 0 must be column 3, day 1 column 4, etc.). Loop stops the column before the last day
    for(j in 3:(ncol(temp)-1)){
      #assign the number of flies that died in the vial on that day (starting with day 1) to the variable 'z'
      z<-temp[1,j]-temp[1,j+1]
        #if >0 flies died add this information to the results dataframe
        if(z>0){
          #iterate over 1 through total number of dead flies
          for(k in 1:z){
            #add a new row to the 'results' data.frame for the given dead fly, specifying vial #, treatment, day died, and
            #record the current vial #
            vial<-temp[,1]
            #record the genotype of the current vial
            treatment<-temp[,2]
            #record the death date of the flies that died on this day (assumes that your input DF starts with day 0 in column 3)
            dd<-j-2
            #append this information into a new row in the 'results' data.frame, and add a '1' in the 4th column to indicate mortality
            results[nrow(results)+1,]<- c(vial,treatment,dd,1)
          } #close for loop
        } #close if loop
    } #close for loop
  
  #now assign the number of flies remaining in the vial on the last day (value in the last column of the row) to the variable 'z'
  z<-temp[1,j+1]
    #if there are any flies alive in the vial on the last day
    if(z>0){
      #iterate over 1:(number of flies alive on the last day)
      for(l in 1:z){
        #record the current vial #
        vial<-temp[,1]
        #record the genotype of the current vial
        treatment<-temp[,2]
        #record the last day we recorded this fly alive (assumes that your input DF starts with day 0 in column 3)
        dd<-j-2
        #append this information into a new row in the 'results' data.frame, and add a '0' in the 4th column to indicate that the fly made it to the end of the experiment
        results[nrow(results)+1,]<- c(vial,treatment,dd,0)
      } #close for loop
    } #close if loop
  } #close original for loop
results$dead<-as.numeric(results$dead)  #reiterate that this column must be class numeric
results$status<-as.numeric(results$status)  #reiterate that this column must be class numeric
results$vial <- as.factor(results$vial) # make sure vial is considered a factor
# gives you only the results dataframe as output from function 
return(results) 
} #close function
```


Read in raw data
```{r}
#read the file from csv
df<-read.csv("~/Desktop/Github/Unckless_Lab_Resources/Infection_survival_analyses/20230509-inn-16Cq/20230509-16Cq-innubila-infection-2.csv")

# only want columns with vial, treatment, and daily mortality
# remove columns not needed for each subset
df<-df[,c(1,4,12:33)]

# remove the last row
# for some reason it is left on there
df <- df[-16,]
```

### Convert each of these dataframes to long and tidy format using function defined above
```{r}
df.convert<-convert_df(df)
```


# plot the survival curve not including confidence intervals

```{r}
# change to not have confidence intervals in this one so you can see them 
df_fit<- survfit(Surv(dead, status) ~ treatment, data=df.convert)
ggsurvplot(df_fit,
          pval = FALSE, conf.int = TRUE,
          #risk.table = TRUE, # Add risk table
          #risk.table.col = "strata", # Change risk table color by groups
          #linetype = "strata", # Change line type by groups
          #surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw(), # Change ggplot2 theme
          palette = c ("aquamarine", "lightcoral")) + ylab("Survival Proporation") + xlab("Days post infection")
```

Get some model summary statistics 
```{r}
# new model using cox proportional hazard (which is basically what above is doing)
df_fit2<- coxph(Surv(dead, status) ~ treatment, data=df.convert)
# look at the statistics of the model
summary(df_fit2)
```

Combine replicates 1 and 2 

```{r}
#read the file from csv
df2<-read.csv("~/Desktop/Github/Unckless_Lab_Resources/Infection_survival_analyses/20230509-inn-16Cq/combined-poked-16Cq-infections.csv")

# only want columns with vial, treatment, and daily mortality
# remove columns not needed for each subset
df2<-df2[,c(1,6,14:34)]

```

### Convert each of these dataframes to long and tidy format using function defined above
```{r}
df2.convert<-convert_df(df2)

# split the column 
df2.convert_S <- str_split_fixed(df2.convert$treatment, "-", 2)

# change column names
colnames(df2.convert_S) <- c("Treatment", "block")

# add columns to df 
df.convert_full <- cbind(df2.convert,df2.convert_S)


```

# plot the survival curve not including confidence intervals

```{r}
# order treatments 
df.convert_full <- df.convert_full%>% 
  mutate(treatment = factor(treatment, levels = c("SP", "16DiNV")))

# change to not have confidence intervals in this one so you can see them 
df2_fit<- survfit(Surv(dead, status) ~ Treatment, data=df.convert_full)
ggsurvplot(df2_fit, size = 5,
          pval = FALSE, conf.int = FALSE,
          ggtheme = theme_light(), # Change ggplot2 theme
          font.tickslab = c(14),
          #legend = "right",
          font.x = c(16),
          font.y = c(16),
          title = "Male D. innubila Poked with 16Cq DiNV",
          legend.title="Treatment",
          legend.labs=c( "16Cq DiNV", "Sterile Poke"),
          font.legend = c(14),
          palette = c("#DF65B0", "#E7E1EF")) + ylab("Survival Proporation") + xlab("Days post infection") 
```

Model with treatment and block considered 

```{r}
# model with just treatment
df_fit_t<- coxph(Surv(dead, status) ~ Treatment, data=df.convert_full)
summary(df_fit_t)

# model with treatment and block 
df_fit_tb<- coxph(Surv(dead, status) ~ Treatment + block, data=df.convert_full)
# look at the statistics of the model
summary(df_fit_tb)

# model with interaction term 
df_fit_tbin<- coxph(Surv(dead, status) ~ Treatment + block + Treatment * block, data=df.convert_full)
# look at the statistics of the model
summary(df_fit_tbin)

# block is not significant, but I will test if I should keep the block in there 

# no block
extractAIC(df_fit_t)
# 568.8226

# with block 
extractAIC(df_fit_tb)
# 569.3603 

# Compare AICs

exp((569.3603 - 568.8226)/2)
# 1.308459
# this would be a p value comparing the block and non-block models. There is no significant difference between the two AICs, so either one is a good fit 
```





