---
title: "20230302-fluid-concentration-qPCR-analysis"
output: github_document
date: '2023-03-03'
---

Load packages 
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```

Read in the data
```{r}
Cq_data <- read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/qPCR_analysis/20230302-fluid-concentration/20230302-fluid-concentration.csv")

```

Histogram of all Cq values 

```{r}

ggplot(Cq_data, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(~sample_type) 

# fill for which primer 
# facet breaks up the graph into different components by sample type
# dodge makes each primer have it's own column per value 

```

Can see obvious differences between the ones that should have virus and the ones that don't. Not as much difference between the non-concentrated and the concentrated but that might be ok. Lambda values are also different, but I wasn't able to add the same amount of DNA to each reaction so it wasn't normalized. 

Calculate variances for the qPCR replicates. We are hoping that the variances are small between the replicates. The bars in the graph above are tightly clustered, and usually have a count of 3 which is a good sign.  But we need to check mathamatically 
```{r}
# make a dummy variable that is just the columns we want
x <- paste(Cq_data$sample_ID, Cq_data$primer, Cq_data$sample_type, Cq_data$date)

# calculate the variances 
# tapply breaks the dataset up into groups and apply a function to each group
# we want each sample_ID/primer/day/sample_type group to have the variance calculated for the Cq value 
# and save this information as a variable 
variances <- tapply(Cq_data$Cq, x, var)

# look at a histogram of the variances 
hist(variances)
# one has really high variance which is not good 

# which variances are above 0.2, which is higher than we want 
variances[which(variances > 0.2)]

```

Most variances are very low. There are only two variances that are above 0.2, and not by much. For both of these if I look at the Cq numbers, it doesn't seem like one Cq is the value making the variance high, they are all a bit different. So I will keep these for now but it's something to know about. 

Now I need to calculate the mean Cq for the replicates

```{r}
# what are the Cq means for each sample_ID/primer/day/sample_type group?
means<- tapply(Cq_data$Cq, x, mean)

# look at the means with a histogram
hist(means)

# want to plot the variances versus the means 
plot(variances~means)

```

There is no real relationship here between what the Cq is and what the variance is. However I noticed here that at least one set of Cqs is not on here 

Where is the sample 13 values?

```{r}
# is it in the variable?
print(x)
#yes 

# is it in the means?
print(means)
# no, because there is 1 NA it seems like it didn't do a mean. Should I remove that one line that has the NA? 
# the other NA is sample 6 where all the Cqs were NA so nothing should come out of that 
```

Going to change my data frame so that it has only one row per sample/primer and has the mean and the variance of the Cq as well 

```{r}
# I know that the variances is an array 
# transform array into df
variances_df <- as.data.frame.table(variances)
# this gives me a dataframe but the entire first column is all the variables 
# and the second column is the variances which it calls freq (not sure why)
# use dplyr to separate them out into separate columns 

# name all the columns I want 
variances_df_sep <- variances_df %>% separate(Var1, c('sample_ID', 'primer', 'sample_type', "date"))
# check dataframe 
head(variances_df_sep)
# rename the Freq column to var for variance
names(variances_df_sep)[5] <- "var"
# check dataframe again
head(variances_df_sep)

```

Basically do the same thing for the means information and add it in 

```{r}
# going to transform the means array into a df
means_df <- as.data.frame.table(means)
# reaname the Freq column to mean
names(means_df)[2] <- "mean"
head(means_df)

# the means and the variances dfs are in the same order which is nice
# now I want to add the mean column to the variances df
means_variances_of_Cq_df <- cbind(variances_df_sep, means = means_df$mean)
# check how it looks
head(means_variances_of_Cq_df)
```


Plot the raw means of Cqs by primer and sample type
```{r}
ggplot(means_variances_of_Cq_df, aes(y= means, x=primer, fill=sample_type)) + geom_boxplot()
```

Flowthrough 115 is not in here because it is NAs in my data. 

Separate out the dataframe to just non-concentrated samples, then calculate the delta Cq between lambda and 115

```{r}
# separate out sample type to non-concentrate
NC_means <- means_variances_of_Cq_df[which(means_variances_of_Cq_df$sample_type == "nonconcentrated"),]

# Separate that dataframe, incriminating by 2, every number between 1-4 
NC_means$means[seq(1,4,2)] # these are the 115 Cq means 
NC_means$means[seq(2,4,2)] # these are lambda primer Cq mean

# make delta Cq, subtract the 115 value from the lambda primer value 
delta_Cqs_NC <- NC_means$means[seq(2,4,2)] - NC_means$means[seq(1,4,2)]

# use barplot to look at the delta Cq for each sample
# use paste to put in the row information, and the seq section to just separate out the rows I want, the 2^ to the delta_Cqs_cells takes it from a negative to a better ratio, and the las=2 makes the labels for the box plots go vertical and readable 
barplot(2^delta_Cqs_NC, names= paste(NC_means$date)[seq(1,4,2)], las=2)
```

I am not sure what this means yet. Should I have not separated out the groups? Maybe I should try to do this looking at all of them at once? 


Calculating Delta Cq for all data 

```{r}
# Separate the means and variances dataframe, incriminating by 2, every number between 1-16 
means_variances_of_Cq_df$means[seq(1,16,2)] # these are the 115 Cq means 
means_variances_of_Cq_df$means[seq(2,16,2)] # these are lambda primer Cq mean

# make delta Cq, subtract the 115 value from the lambda primer value 
delta_Cqs <- means_variances_of_Cq_df$means[seq(2,16,2)] - means_variances_of_Cq_df$means[seq(1,16,2)]

# use barplot to look at the delta Cq for each sample
# use paste to put in the row information, and the seq section to just separate out the rows I want, the 2^ to the delta_Cqs_cells takes it from a negative to a better ratio, and the las=2 makes the labels for the box plots go vertical and readable 
# increase margin size
par(mar = c(11, 3, 3, 3))
barplot(2^delta_Cqs, names= paste(means_variances_of_Cq_df$date, means_variances_of_Cq_df$sample_type)[seq(1,16,2)], las=2,)
```




