---
title: "20230307-Myd88-Dv-1-qPCR-DiNV-Analysis"
output: github_document
date: '2023-03-07'
---

Load packages needed 

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```

Load in the two datasets, supernatant and cells:

```{r}
supernatant_Cqs <- read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/qPCR_analysis/20230307-Myd88-Dv-1/20230307-supernatants-Cqs.csv")

cells_Cqs <- read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/qPCR_analysis/20230307-Myd88-Dv-1/20230307-cells-Cqs.csv")
```

Histogram of all Cq values for Supernatant

```{r}
# supernatant
ggplot(supernatant_Cqs, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(cell_type~day) 
# fill for which primer 
# facet breaks up the graph into different components 
# dodge makes each primer have it's own column per value 
```

The NA for day is the 40ul original aliquot/innoculum sample. 

Histogram of all Cq values for Cells

```{r}
# cells
ggplot(cells_Cqs, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(cell_type~day) 
# fill for which primer 
# facet breaks up the graph into different components 
# dodge makes each primer have it's own column per value 
```

You can already tell that it seems like nothing is going on with DiNV for the Myd88 cells in terms of replication. 

Calculate the variances between the qPCR replicates for supernatant samples
```{r}
# make a dummy variable that is just the columns we want
s <- paste(supernatant_Cqs$sample_ID, supernatant_Cqs$primer, supernatant_Cqs$cell_type, supernatant_Cqs$day, supernatant_Cqs$infection)

# calculate the variances 
# tapply breaks the dataset up into groups and apply a function to each group
# we want each sample_ID/primer/day/cell_type group to have the variance calculated for the Cq value 
# and save this information as a variable 
sup_variances <- tapply(supernatant_Cqs$Cq, s, var)

# look at a histogram of the variances 
hist(sup_variances)
# one has really high variance which is not good 

# which variances are above 0.2, which is higher than we want 
sup_variances[which(sup_variances > 0.2)]

```

Ok looks like there are 2 variances that are too high, both for non-infected 115 reactions. That could make sense because there is nothing to amplify so it is getting not good signal. I am not sure if there is a replicate to remove in these that would make the variances go down meaningfully. Right now I will leave them. 


Calculate the variances between the qPCR replicates for cell samples
```{r}
# make a dummy variable that is just the columns we want
c <- paste(cells_Cqs$sample_ID, cells_Cqs$primer, cells_Cqs$cell_type, cells_Cqs$day, cells_Cqs$infected.)

# calculate the variances 
# tapply breaks the dataset up into groups and apply a function to each group
# we want each sample_ID/primer/day/cell_type group to have the variance calculated for the Cq value 
# and save this information as a variable 
cell_variances <- tapply(cells_Cqs$Cq, c, var)

# look at a histogram of the variances 
hist(cell_variances)
# one has really high variance which is not good 

# which variances are above 0.2, which is higher than we want 
cell_variances[which(cell_variances > 0.2)]

```

Again, I am not sure if I should remove any of the replicates. For these, there does seem to be one replicate that is somewhat off. But, the overall variances are not too worrying. For now I will leave them?

Calculate the mean Cq values for supernatants 

```{r}
# what are the Cq means for each sample_ID/primer/day/sample_type group?
sup_means <- tapply(supernatant_Cqs$Cq, s, mean, na.rm=T)

# look at the means with a histogram
hist(sup_means)

# want to plot the variances versus the means 
plot(sup_variances~sup_means)

```

This shows that as the Cq goes up/approaches the limit, the variance increases. This sort of makes sense, that the values get more erratic at the edge of detection. 

Calculate the mean Cq values for Cells

```{r}
# what are the Cq means for each sample_ID/primer/day/sample_type group?
cell_means <- tapply(cells_Cqs$Cq, c, mean, na.rm=T)

# look at the means with a histogram
hist(cell_means)

# want to plot the variances versus the means 
plot(cell_variances~cell_means)

```

Again, there is the same pattern here of increasing variance with increased Cq mean. However it is not just the close to 40 Cqs, but the "higher" variances are much smaller than those for the supernatant. An interesting pattern. I'm not sure what to make of it. 

Now I need to add the mean and variance values to my main dataframes. 
First I will do this for the supernatant 

```{r}
# transform array into df
sup_variances_df <- as.data.frame.table(sup_variances)
# this gives me a dataframe but the entire first column is all the variables 
# and the second column is the variances which it calls freq (not sure why)
# use dplyr to separate them out into separate columns 

# name all the columns I want 
sup_variances_df_sep <- sup_variances_df %>% separate(Var1, c('sample_ID', 'primer', 'cell_type', 'day', 'infected'))
# check dataframe 
head(sup_variances_df_sep)
# rename the Freq column to var for variance
names(sup_variances_df_sep)[6] <- "var"
# check dataframe again
head(sup_variances_df_sep)
```

So I have a df with the variances, now I want the means in it as well 
```{r}
# going to transform the means array into a df
sup_means_df <- as.data.frame.table(sup_means)
# reaname the Freq column to mean
names(sup_means_df)[2] <- "mean"
head(sup_means_df)

# the means and the variances dfs are in the same order which is nice
# now I want to add the mean column to the variances df
sup_means_variances_of_Cq_df <- cbind(sup_variances_df_sep, means = sup_means_df$mean)
# check how it looks
head(sup_means_variances_of_Cq_df)
```

Do this same process for the cells data 

```{r}
# transform array into df
cells_variances_df <- as.data.frame.table(cell_variances)
# this gives me a dataframe but the entire first column is all the variables 
# and the second column is the variances which it calls freq (not sure why)
# use dplyr to separate them out into separate columns 

# name all the columns I want 
cells_variances_df_sep <- cells_variances_df %>% separate(Var1, c('sample_ID', 'primer', 'cell_type', 'day', 'infected'))
# check dataframe 
head(cells_variances_df_sep)
# rename the Freq column to var for variance
names(cells_variances_df_sep)[6] <- "var"
# check dataframe again
head(cells_variances_df_sep)
```

So I have a df with the variances, now I want the means in it as well 
```{r}
# going to transform the means array into a df
cell_means_df <- as.data.frame.table(cell_means)
# reaname the Freq column to mean
names(cell_means_df)[2] <- "mean"
head(cell_means_df)

# the means and the variances dfs are in the same order which is nice
# now I want to add the mean column to the variances df
cell_means_variances_of_Cq_df <- cbind(cells_variances_df_sep, means = cell_means_df$mean)
# check how it looks
head(cell_means_variances_of_Cq_df)
```

Plot the raw means of Cqs by primer for supernatant by day 
```{r}
ggplot(sup_means_variances_of_Cq_df, aes(y= means, x=primer, fill=infected)) + geom_boxplot() + facet_grid(~day) 
```

Plot the raw means of Cqs by primer for cells by day 
```{r}
ggplot(cell_means_variances_of_Cq_df, aes(y= means, x=primer, fill=infected)) + geom_boxplot() + facet_grid(~day) 
```

Starting with just supernatant, calculate the delta Cq, then visualize 

```{r}
# separate out the 40ul sample because that one won't fit in 
sup_means_var <- sup_means_variances_of_Cq_df[which(sup_means_variances_of_Cq_df$cell_type != "innubila"),]

# Separate that dataframe, incriminating by 2, every number between 1-30 (number of rows in dataframe)
sup_means_var$means[seq(1,30,2)] # these are the 115 Cq means 
sup_means_var$means[seq(2,30,2)] # these are the lambda primer Cq means 

# make delta Cq, subtract the 115 value from the lambda primer value 
delta_Cqs_supernatants <- sup_means_var$means[seq(2,30,2)] - sup_means_var$means[seq(1,30,2)]

# use barplot to look at the delta Cq for each sample
# use paste to put in the row information, and the seq section to just separate out the rows I want, the 2^ to the delta_Cqs_cells takes it from a negative to a better ratio, and the las=2 makes the labels for the box plots go vertical and readable 
par(mar = c(8, 3, 3, 3))
barplot(2^delta_Cqs_supernatants, names= paste(sup_means_var$cell_type, sup_means_var$day, sup_means_var$infected)[seq(1,30,2)], las=2)

```

I wish I knew a way to change the order of these bar plots. 


Doing cells, calculate the delta Cq, then visualize 

```{r}

# Separate that dataframe, incriminating by 2, every number between 1-32 (number of rows in dataframe)
cell_means_variances_of_Cq_df$means[seq(1,32,2)] # these are the 115 Cq means 
cell_means_variances_of_Cq_df$means[seq(2,32,2)] # these are the RPL11 or RP49 primer Cq means 

# make delta Cq, subtract the 115 value from the control primer value 
delta_Cqs_cells <- cell_means_variances_of_Cq_df$means[seq(2,32,2)] - cell_means_variances_of_Cq_df$means[seq(1,32,2)]

# use barplot to look at the delta Cq for each sample
# use paste to put in the row information, and the seq section to just separate out the rows I want, the 2^ to the delta_Cqs_cells takes it from a negative to a better ratio, and the las=2 makes the labels for the box plots go vertical and readable 
par(mar = c(8, 3, 3, 3))
barplot(2^delta_Cqs_cells, names= paste(cell_means_variances_of_Cq_df$cell_type, cell_means_variances_of_Cq_df$day, cell_means_variances_of_Cq_df$infected)[seq(1,32,2)], las=2)
```

Ok this kind of makes sense, it basically seems to show nothing for Myd88, but stuff for Dv-1. This is interesting when you compare it to the regular PCR data..  

Should I be making delta Cq calculations with this? 

What is the delta Cq for the day 26 innubila primary fluid sample?

```{r}
mean115 <- 24.11333
meanlambda <- 24.16333
deltaCq <- meanlambda - mean115
print(deltaCq)
deltaCq2 <- 2^deltaCq
print(deltaCq2)
```


Can I remake the delta Cq plots in a better format? 
Start with cells 
```{r}
# what format is this 
delta_Cqs_cells
# ok it is just a list of values of the delta Cq 
# I put this into a csv file 

D_Cq_cells <- read.csv("delta_cq_cells.csv")

#reorder days 
results_factor_levels <- c("5", "0")
# then apply this to the CCM data
D_Cq_cells$day <- factor(D_Cq_cells$day, levels=results_factor_levels)

ggplot(D_Cq_cells, aes(y= Delta_Cq_2, x=cell, fill=infection_status)) + geom_boxplot() + facet_grid(~day) + theme_linedraw() 

ggsave("delta_Cq_cells.png")

```

Now do supernatant
```{r}
# what format is this 
delta_Cqs_supernatants
# ok it is just a list of values of the delta Cq 
# I put this into a csv file 

D_Cq_sup <- read.csv("delta_cq_sup.csv")

#reorder days 
results_factor_levels <- c("5", "0")
# then apply this to the CCM data
D_Cq_sup$day <- factor(D_Cq_sup$day, levels=results_factor_levels)

ggplot(D_Cq_sup, aes(y= Delta_Cq_2, x=cell, fill=infection_status)) + geom_boxplot() + facet_grid(~day) + theme_linedraw()

ggsave("delta_Cq_sup.png")

```

How might I do delta delta Cq? 

Cells

```{r}
# I calculated the delta delta Cq by hand 


delta_delta <- read.csv("delta_delta_Cq_hand.csv")

# make all NAs equal to 1 because the relative increase in virus in just 1, no increase 

delta_delta <- replace(delta_delta, is.na(delta_delta), 1)

# want to force axis to top out at 5 
ggplot(delta_delta, aes(y= X2_to_the_delta_delta, x=cell_type, color=infected)) + geom_point(position="jitter") + facet_grid(~sample_type) + theme_linedraw() + ylim(0, 5)



```

Come back to this analysis for supernatant and try to plot just raw Cqs for 115


Lets look at just cell Cqs - but using the means of the replicates
```{r}
# separate out just 115 
cell_means_variances_of_Cq_df
raw_115_cells <- cell_means_variances_of_Cq_df[which(cell_means_variances_of_Cq_df$primer == "115"),]

# make NaN as 40 
raw_115_cells <- replace(raw_115_cells, is.na(raw_115_cells), 40)

ggplot(raw_115_cells, aes(y= means, x=day, color=infected)) + geom_point(position="jitter") + facet_grid(~cell_type) + theme_linedraw()

```

Supernatant Cqs - but using the means of the replicates
```{r}
# separate out just 115 
sup_means_variances_of_Cq_df
# remove the innubila line 
raw_115_sup <- sup_means_variances_of_Cq_df[which(sup_means_variances_of_Cq_df$cell_type != "innubila"),]
# make it just 115
raw_115_sup <- raw_115_sup[which(raw_115_sup$primer == "115"),]

# make NaN as 0 
raw_115_sup <- replace(raw_115_sup, is.na(raw_115_sup), 40)

ggplot(raw_115_sup, aes(y= means, x=day, color=infected)) + geom_point(position="jitter") + facet_grid(~cell_type) + theme_linedraw()

```












