---
title: "20240112-exo-treatment-test"
format: gfm
editor: visual
---

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width=7)
```

Load in dataset

```{r}
Cq_values <- read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/qPCR_analysis/20240112-exo-treatment-test/20240112-exo-qPCR.csv")
# I only need rows 1 - 48 
Cq_values <- Cq_values[c(1:48),]

```

Look at raw Cq values by primer and sample type

```{r}
# if you want to also include another varaible to separate out the histogram by, you can include a facet 
ggplot(Cq_values, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(~type) 
```

Based off of this, it looks like the exonuclease treatment may have done something, the TPI Cqs are higher... but also the PIF3

Calculate variance and mean of Cq value between the technical replicates

```{r}
# use the variance function, and calculates the variance in Cq by the unique.name (each sample/primer has 3 Cq values to 
# calculate the variance by)
Cq_values$Cq_var <- ave(Cq_values$Cq, Cq_values$unique_name, FUN=var)

# use the mean function, and calculates the mean of Cq values by the unique.name (each sample/primer has 3 Cq values to 
# calculate the mean by)
Cq_values$Cq_mean <- ave(Cq_values$Cq, Cq_values$unique_name, FUN=mean)

# Keep all rows where the replicate is 1 (or you could do 2 or 3)
# make into new Df so we keep the original with all the Cq values
Cq_values_1rep <- Cq_values[which(Cq_values$replicate == "1"),]

# histogram of all variances
ggplot(Cq_values_1rep, aes(x=Cq_var)) + geom_histogram(bins = 50)

# all of these seem pretty low, there is maybe two that are a little high, but not much 
# right now I am going to leave them

# plot variances and means as a scatterplot 
ggplot(Cq_values_1rep, aes(x=Cq_mean, y=Cq_var)) +
  geom_point(size=2, shape=23)


```

The only slightly higher variance samples have the highest Cq, which is what we have seen before.

Calculating Delta Cqs between TPI and PIF3

```{r}
# samples are not ordered properly
# they need to be ordered with the two primer samples next to each other for each sample 
# order them by the sample
Cq_values_1rep <- Cq_values_1rep[order(Cq_values_1rep$sample),]
Cq_values_1rep
nrow(Cq_values_1rep)

# Separate that dataframe, incriminating by 2, every number between 1-16 (number of rows in dataframe)
Cq_values_1rep$Cq_mean[seq(1,16,2)] # these are the TPI Cq means 
Cq_values_1rep$Cq_mean[seq(2,16,2)] # these are the PIF 3 primer Cq means 

# make the delta Cq by subtracting the PIF 3 values from the TPI primer values
# and this is saved as a vector in R 
delta_Cqs <- Cq_values_1rep$Cq_mean[seq(1,16,2)] - Cq_values_1rep$Cq_mean[seq(2,16,2)]
#vector
delta_Cqs

# Keep only rows that are PIF3 because now we have two rows per sample
Cq_values1rep_Delta <- Cq_values_1rep[which(Cq_values_1rep$primer == "PIF3"),]
# And then add in the delta Cqs as a new column
Cq_values1rep_Delta$delta_Cq <- delta_Cqs

# do 2^ delta Cq
Cq_values1rep_Delta$delta_Cq_2 <- 2^(delta_Cqs)

# plot 
ggplot(Cq_values1rep_Delta, aes(y= delta_Cq_2, x=type, color=sample_type)) + geom_boxplot()  + theme_linedraw() + geom_point(position="jitter", size=3) 

```

It looks from this that the exonuclease treatment does make a difference, and the largest difference is with the cell pellet samples. Although the clarification pellet ones still have a better ratio of virus to cell than the non-treated ones.
