---
title: "20240221-low-dilutions-rep-2-male"
format: gfm
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width=11, fig.height = 11)
```

Load in packages needed for the analysis

```{r}
library("survival")
library("survminer")
library(dplyr)
library(tidyr)
library(stringr)
```

### Loop to convert the example data.frame 'df' into properly formatted data.frame 'results'

```{r}
#write a function to transform a data.frame that has the column format 'vial | treatment | D0 | D1 | D2...', with one row for each vial
#into a long version in tidy format that can be input to make a survivorship curve
convert_df<-function(df){
  #open empty data.frame to store results
  results<-data.frame(vial=character(),treatment=character(),dead=numeric(),status=numeric())
  #This loop will index out each row (one row per vial) one at a time, transform it into long format (one row per fly), and add the information to the empty data.frame called results
  for(i in 1:nrow(df)){
  #isolate the row (vial) you want to work on
  temp<-df[i,]
    #iteratively isolate each day for this vial (day 0 must be column 3, day 1 column 4, etc.). Loop stops the column before the last day
    for(j in 3:(ncol(temp)-1)){
      #assign the number of flies that died in the vial on that day (starting with day 1) to the variable 'z'
      z<-temp[1,j]-temp[1,j+1]
        #if >0 flies died add this information to the results dataframe
        if(z>0){
          #iterate over 1 through total number of dead flies
          for(k in 1:z){
            #add a new row to the 'results' data.frame for the given dead fly, specifying vial #, treatment, day died, and
            #record the current vial #
            vial<-temp[,1]
            #record the genotype of the current vial
            treatment<-temp[,2]
            #record the death date of the flies that died on this day (assumes that your input DF starts with day 0 in column 3)
            dd<-j-2
            #append this information into a new row in the 'results' data.frame, and add a '1' in the 4th column to indicate mortality
            results[nrow(results)+1,]<- c(vial,treatment,dd,1)
          } #close for loop
        } #close if loop
    } #close for loop
  
  #now assign the number of flies remaining in the vial on the last day (value in the last column of the row) to the variable 'z'
  z<-temp[1,j+1]
    #if there are any flies alive in the vial on the last day
    if(z>0){
      #iterate over 1:(number of flies alive on the last day)
      for(l in 1:z){
        #record the current vial #
        vial<-temp[,1]
        #record the genotype of the current vial
        treatment<-temp[,2]
        #record the last day we recorded this fly alive (assumes that your input DF starts with day 0 in column 3)
        dd<-j-2
        #append this information into a new row in the 'results' data.frame, and add a '0' in the 4th column to indicate that the fly made it to the end of the experiment
        results[nrow(results)+1,]<- c(vial,treatment,dd,0)
      } #close for loop
    } #close if loop
  } #close original for loop
results$dead<-as.numeric(results$dead)  #reiterate that this column must be class numeric
results$status<-as.numeric(results$status)  #reiterate that this column must be class numeric
results$vial <- as.factor(results$vial) # make sure vial is considered a factor
# gives you only the results dataframe as output from function 
return(results) 
} #close function
```

Read in raw data

**Note that for these datasets, the adjusted N number is the number of flies alive on day 2 because there seemed to be more early death in these experiments than usual**

```{r}
#read the file from csv
df<-read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/Infection_survival_analyses/20240221-low-dilutions-rep-2-male/20240221-male-low-dil-rep2.csv")

# now extra rows, I already separated the females out 
# separate out columns needed
df <- df[,c(1,9,14:28)]

```

Convert dataframe

```{r}
df.convert<-convert_df(df)
```

Plot survivial curve with every line separate

```{r}
# change to not have confidence intervals in this one so you can see them 
df_fit<- survfit(Surv(dead, status) ~ treatment, data=df.convert)
ggsurvplot(df_fit, size = 5,
          pval = FALSE, conf.int = FALSE,
          font.tickslab = c(14),
          font.x = c(16),
          font.y = c(16),
          ggtheme = theme_light())
          

```

**Combine Replicates 1 and 2**

```{r}
#read the file from csv
df2<-read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/Infection_survival_analyses/20240221-low-dilutions-rep-2-male/low-dil-male-combo.csv")

# now extra rows, I already separated the females out 
# separate out columns needed
df2 <- df2[,c(1,8,13:27)]

```

Convert dataframe

```{r}
df2.convert<-convert_df(df2)
```

Plot survivial curve with every line separate

```{r}
# level the treatments so they go in an order 
df2.convert <- df2.convert %>% 
  mutate(treatment = factor(treatment, levels = c("CCM", "0.005 FFU", "0.01 FFU", "0.05 FFU")))
# I'm not sure this worked

# change to not have confidence intervals in this one so you can see them 
# for some reason when I render this plot it does not do the colors in the order I see them in R. I'm not sure why 
df2_fit<- survfit(Surv(dead, status) ~ treatment, data=df2.convert)
ggsurvplot(df_fit, size = 5,
          pval = FALSE, conf.int = FALSE,
          legend = "right",
          font.tickslab = c(14),
          font.x = c(16),
          font.y = c(16),
          ggtheme = theme_light(),
          title = "Male D. innubila Injected with Dilutions of Passage 4 DiNV",
          legend.title="Treatment",
          legend.labs=c(  "0.01 FFU", "0.005 FFU", "0.05 FFU", "CCM"),
          font.legend = c(14),
          palette = c(  "#E7298A","#C994C7", "#980043", "#E7E1EF")) + ylab("Survival Proporation") + xlab("Days post injection")

```

Find median survival time by treatment

```{r}
surv_median(df2_fit, combine = FALSE)
```

Model just looking at significance of block and treatment

```{r}
# add in block 
df2.convert$Block <- rep(c("A","B"), c(77, 73))
# model including block 
df3_fit<- coxph(Surv(dead, status) ~ treatment + Block, data=df2.convert)
summary(df3_fit)

```

Look at model using dilution as a continuous variable

```{r}
# make dilution as a numeric/continuous variable
# make a duplicate column 
df2.convert$numeric.dilution <- df2.convert$treatment
# set CCM to a dose of 0
df2.convert$numeric.dilution <- str_replace_all(df2.convert$numeric.dilution, 'CCM', '0')
# remove FFU and make numeric
df2.convert$numeric.dilution <- as.numeric(gsub("FFU", "", df2.convert$numeric.dilution))

# model including block 
df4_fit<- coxph(Surv(dead, status) ~ numeric.dilution + Block, data=df2.convert)
summary(df4_fit)

```

Only look at 0.05 and 0.01

```{r}
df2.convert_51 <- df2.convert[which(df2.convert$numeric.dilution != 0),]
df2.convert_51 <- df2.convert_51[which(df2.convert_51$numeric.dilution != 0.005),]

df5_fit<- coxph(Surv(dead, status) ~ numeric.dilution + Block, data=df2.convert_51)
summary(df5_fit)
```

Only look at and 0.01 and .005

```{r}
df2.convert_051 <- df2.convert[which(df2.convert$numeric.dilution != 0),]
df2.convert_051 <- df2.convert_051[which(df2.convert_051$numeric.dilution != 0.050),]

df5_fit<- coxph(Surv(dead, status) ~ numeric.dilution + Block, data=df2.convert_051)
summary(df5_fit)
```

Only look at and CCM and .005

```{r}
df2.convert_05C <- df2.convert[which(df2.convert$numeric.dilution != 0.010),]
df2.convert_05C <- df2.convert_05C[which(df2.convert_05C$numeric.dilution != 0.050),]

df6_fit<- coxph(Surv(dead, status) ~ numeric.dilution + Block, data=df2.convert_05C)
summary(df6_fit)
```
