---
title: "p4-dilutions-time"
format: gfm
editor: visual
---

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(Rmisc)
library(ggpubr)
library(rstatix)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width=7, fig.height = 6)
```

Load in dataset

```{r}
Cq_values <- read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/qPCR_analysis/20240228-viral-load-dilutions/viral-dil-time-sheet.csv")
```

Separate male and female for now

```{r}
# males
Cq_values_M <- Cq_values[which(Cq_values$sex == "male"),]
# females
Cq_values_F <- Cq_values[which(Cq_values$sex == "female"),]
```

**Female Analysis First**

Look at raw Cq values

```{r}
ggplot(Cq_values_F, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(~dilution) 
```

TPI looks remarkably consistant, PIF3 is all over the place but this isn't separated out by day which would make differences.

Look at raw Cq values and also separate by day

```{r}
ggplot(Cq_values_F, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(dilution ~ day) 
```

This is showing more patterns that we sort of expect, but we have to do the delta Cq comparison to the TPI primer to be sure

Calculating female delta Cq 

```{r}
# need to organize by name of the sample I think
Cq_values_F <- Cq_values_F[order(Cq_values_F$sample_ID),]
# this worked to order by the sample number 
# and the PIF3 value is first in the order and TPI second

# Separate that dataframe, incriminating by 2, every number between 1-192 (number of rows in dataframe)
Cq_values_F$Cq[seq(1,192,2)] # these are the PIF 3 Cq #s
Cq_values_F$Cq[seq(2,192,2)] # these are the TPI primer Cq #s 

# make the delta Cq by subtracting the PIF 3 values from the TPI primer values
# and this is saved as a vector in R 
delta_Cqs_F <- Cq_values_F$Cq[seq(2,192,2)] - Cq_values_F$Cq[seq(1,192,2)]
#vector
delta_Cqs_F

# Make a new dataframe that only has one row per sample by getting rid of the rows with the TPI primer
Cq_values_F_Delta <- Cq_values_F[which(Cq_values_F$primer == "PIF3"),]

# And then add in the delta Cqs as a new column
Cq_values_F_Delta$delta_Cq <- delta_Cqs_F

# add a column with 2^ delta Cq
Cq_values_F_Delta$delta_Cq_2 <- 2^(delta_Cqs_F)

```
Female
Doing t test statistics and plotting 
```{r}
# doing pairwise t-tests between dilution methods, within each day
# for delta Cq
fem0.stat.test <- Cq_values_F_Delta %>%
  group_by(day) %>%
  t_test(delta_Cq_2 ~ dilution) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
# add the bracket location info for each comparison
fem0.stat.test<-fem0.stat.test %>% add_xy_position(x = "day")

#do paiwrise t-test between days within dilution methods
 # for delta Cq
fem0.stat.test.2 <- Cq_values_F_Delta %>%
  group_by(dilution) %>%
  t_test(delta_Cq_2 ~ day) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
#add bracket location info
fem0.stat.test.2<-fem0.stat.test.2 %>% add_xy_position(x = "day", group="dilution")

# show statistics 
fem0.stat.test
fem0.stat.test.2

ledgend_title <- "Virus Delivery"

# plot and use a log 10 scale for the y axis 
ggplot(Cq_values_F_Delta, aes(y= delta_Cq_2, x=day)) +
  geom_boxplot(outlier.shape=NA, aes(fill=dilution))  + 
  theme_light() + scale_fill_manual(ledgend_title, values=c("#E7E1EF", "#C994C7", "#CE1256")) + 
  geom_dotplot(aes(fill=dilution), binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=6), legend.title=element_text(size=7)) +
  scale_x_discrete(labels=c("day0" = "0 days", "day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) + 
  labs(title = "Comparing Viral Load in Female D. innubila Across Early Infection",y = "Relative amount of DiNV genome to host genome", x = "Days Post Injection") +
  theme(legend.position = c(0.93, 0.15), legend.background = element_rect(linetype="solid", colour ="black")) +
  stat_pvalue_manual(fem0.stat.test, y.position = seq(3.5,4, by=1.5/3),  hide.ns = TRUE) +
  stat_pvalue_manual(fem0.stat.test.2, y.position = seq(5,10, by=5/7), hide.ns = TRUE)

```

**Male Analysis Second**

Look at raw Cq values and also separate by day

```{r}
ggplot(Cq_values_M, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(dilution ~ day) 
```

Looks pretty similar to the female results

Calculating male delta Cq 

```{r}
# need to organize by name of the sample I think
Cq_values_M <- Cq_values_M[order(Cq_values_M$sample_ID),]
# this worked to order by the sample number 
# and the PIF3 value is first in the order and TPI second

# notice that there is one NA in the data that I forgot to set as 40 for the Cq 

Cq_values_M[is.na(Cq_values_M)] <- 40

# Separate that dataframe, incriminating by 2, every number between 1-192 (number of rows in dataframe)
Cq_values_M$Cq[seq(1,192,2)] # these are the PIF 3 Cq #s
Cq_values_M$Cq[seq(2,192,2)] # these are the TPI primer Cq #s 

# make the delta Cq by subtracting the PIF 3 values from the TPI primer values
# and this is saved as a vector in R 
delta_Cqs_M <- Cq_values_M$Cq[seq(2,192,2)] - Cq_values_M$Cq[seq(1,192,2)]
#vector
delta_Cqs_M

# Make a new dataframe that only has one row per sample by getting rid of the rows with the TPI primer
Cq_values_M_Delta <- Cq_values_M[which(Cq_values_M$primer == "PIF3"),]

# And then add in the delta Cqs as a new column
Cq_values_M_Delta$delta_Cq <- delta_Cqs_M

# add a column with 2^ delta Cq
Cq_values_M_Delta$delta_Cq_2 <- 2^(delta_Cqs_M)

```

Doing t test statistics and plotting 
```{r}
# doing pairwise t-tests between dilution methods, within each day
# for delta Cq
male0.stat.test <- Cq_values_M_Delta %>%
  group_by(day) %>%
  t_test(delta_Cq_2 ~ dilution) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
# add the bracket location info for each comparison
male0.stat.test <- male0.stat.test %>% add_xy_position(x = "day")

# do paiwrise t-test between days within dilution methods
 # for delta Cq
male0.stat.test.2 <- Cq_values_M_Delta %>%
  group_by(dilution) %>%
  t_test(delta_Cq_2 ~ day) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
# add bracket location info
male0.stat.test.2 <- male0.stat.test.2 %>% add_xy_position(x = "day", group="dilution")

# show statistics 
male0.stat.test
male0.stat.test.2

ledgend_title <- "Virus Delivery"

# plot and use a log 10 scale for the y axis 
ggplot(Cq_values_M_Delta, aes(y= delta_Cq_2, x=day)) +
  geom_boxplot(outlier.shape=NA, aes(fill=dilution))  + 
  theme_light() + scale_fill_manual(ledgend_title, values=c("#E7E1EF", "#C994C7", "#CE1256")) + 
  geom_dotplot(aes(fill=dilution), binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=6), legend.title=element_text(size=7)) +
  scale_x_discrete(labels=c("day0" = "0 days", "day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) + 
  labs(title = "Comparing Viral Load in Male D. innubila Across Early Infection",y = "Relative amount of DiNV genome to host genome", x = "Days Post Injection") +
  theme(legend.position = c(0.93, 0.15), legend.background = element_rect(linetype="solid", colour ="black")) +
  stat_pvalue_manual(male0.stat.test, y.position = seq(3.5,4, by=1.5/4),  hide.ns = TRUE) +
  stat_pvalue_manual(male0.stat.test.2, y.position = seq(5,9, by=4/8), hide.ns = TRUE)

```

**Delta Delta Analysis**

Females

```{r}
# this might be a little mess to do 
# separate out the day 0 samples to start with 
F_0 <- Cq_values_F_Delta[which(Cq_values_F_Delta$day == "day0"),]
# separate out dilutions?
F_0_3 <- F_0[which(F_0$dilution == "3 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(F_0_3$delta_Cq)
# -9.25875
F_0_01 <- F_0[which(F_0$dilution == "0.1 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(F_0_01$delta_Cq)
# -13.38625
F_0_001 <- F_0[which(F_0$dilution == "0.01 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(F_0_001$delta_Cq)
# -15.4525

# main data sheet without day 0 
F <- Cq_values_F_Delta[which(Cq_values_F_Delta$day != "day0"),]
# separate out just dilution 3 
F_3 <- F[which(F$dilution == "3 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_3 <- F_3$delta_Cq - -9.25875
delta_delta_3

# add as column 
F_3$delta_delta_Cq <- delta_delta_3

# do 2^ delta delta 
F_3$delta_delta_Cq_2 <- 2^(delta_delta_3)

# separate out just dilution 0.1 
F_01 <- F[which(F$dilution == "0.1 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_01 <- F_01$delta_Cq - -13.38625
delta_delta_01

# add as column 
F_01$delta_delta_Cq <- delta_delta_01

# do 2^ delta delta 
F_01$delta_delta_Cq_2 <- 2^(delta_delta_01)

# separate out just dilution 0.01 
F_001 <- F[which(F$dilution == "0.01 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_001 <- F_001$delta_Cq - -15.4525
delta_delta_001

# add as column 
F_001$delta_delta_Cq <- delta_delta_001

# do 2^ delta delta 
F_001$delta_delta_Cq_2 <- 2^(delta_delta_001)

# add all of the days back together 
Female_delta_delta <- rbind(F_001, F_01, F_3)

```


Statistics and plotting females 
```{r}
# doing pairwise t-tests between dilution methods, within each day
fem.stat.test <- Female_delta_delta %>%
  group_by(day) %>%
  t_test(delta_delta_Cq_2 ~ dilution) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
# add the bracket location info for each comparison
fem.stat.test <- fem.stat.test %>% add_xy_position(x = "day")

# do paiwrise t-test between days within dilution methods
fem.stat.test.2 <- Female_delta_delta %>%
  group_by(dilution) %>%
  t_test(delta_delta_Cq_2 ~ day) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
# add bracket location info
fem.stat.test.2 <- fem.stat.test.2 %>% add_xy_position(x = "day", group="dilution")

# show statistics 
fem.stat.test
fem.stat.test.2

# plot the boxplot with all pairwise stats visualized on top
legend_title <- "Virus Delivery"
ggplot(Female_delta_delta, aes(y= delta_delta_Cq_2, x=day)) + geom_boxplot(outlier.shape=NA,aes(fill=dilution)) +  
  scale_fill_manual(legend_title, values=c("#E7E1EF", "#C994C7", "#CE1256")) + 
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8),aes(fill=dilution)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=10), legend.title=element_text(size=11)) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  labs(title = "Comparing DiNV Increase in Female Flies \nInjected with Various Titers Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection") +
  theme(legend.position = c(0.9, 0.2), legend.background = element_rect(linetype="solid", colour ="black")) +
  stat_pvalue_manual(fem.stat.test, y.position = 8, hide.ns = TRUE) +
  stat_pvalue_manual(fem.stat.test.2, y.position = seq(9,12, by=3/3), hide.ns = TRUE)

```




Males
Calculate delta delta Cq
```{r}
# this might be a little mess to do 
# separate out the day 0 samples to start with 
M_0 <- Cq_values_M_Delta[which(Cq_values_M_Delta$day == "day0"),]
# separate out dilutions?
M_0_3 <- M_0[which(M_0$dilution == "3 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(M_0_3$delta_Cq)
# -9.4875
M_0_01 <- M_0[which(M_0$dilution == "0.1 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(M_0_01$delta_Cq)
# -12.63125
M_0_001 <- M_0[which(M_0$dilution == "0.01 FFU"),]
# find mean of Delta Cq (not 2^delta)
mean(M_0_001$delta_Cq)
# -13.40625

# main data sheet without day 0 
M <- Cq_values_M_Delta[which(Cq_values_M_Delta$day != "day0"),]
# separate out just dilution 3 
M_3 <- M[which(M$dilution == "3 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_3_M <- M_3$delta_Cq - -9.4875
delta_delta_3_M

# add as column 
M_3$delta_delta_Cq <- delta_delta_3_M

# do 2^ delta delta 
M_3$delta_delta_Cq_2 <- 2^(delta_delta_3_M)

# separate out just dilution 0.1 
M_01 <- M[which(M$dilution == "0.1 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_01_M <- M_01$delta_Cq - -12.63125
delta_delta_01_M

# add as column 
M_01$delta_delta_Cq <- delta_delta_01_M

# do 2^ delta delta 
M_01$delta_delta_Cq_2 <- 2^(delta_delta_01_M)

# separate out just dilution 0.01 
M_001 <- M[which(M$dilution == "0.01 FFU"),]

# subtract average day 0 delta Cq for 3 FFU from the delta Cq of all others 
delta_delta_001_M <- M_001$delta_Cq - -13.40625
delta_delta_001_M

# add as column 
M_001$delta_delta_Cq <- delta_delta_001_M

# do 2^ delta delta 
M_001$delta_delta_Cq_2 <- 2^(delta_delta_001_M)

# add all of the days back together 
Male_delta_delta <- rbind(M_001, M_01, M_3)
```

Statistics and plotting males 
```{r}
# doing pairwise t-tests between dilution methods, within each day
male.stat.test <- Male_delta_delta %>%
  group_by(day) %>%
  t_test(delta_delta_Cq_2 ~ dilution) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
# add the bracket location info for each comparison
male.stat.test <- male.stat.test %>% add_xy_position(x = "day")

# do paiwrise t-test between days within dilution methods
male.stat.test.2 <- Male_delta_delta %>%
  group_by(dilution) %>%
  t_test(delta_delta_Cq_2 ~ day) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
# add bracket location info
male.stat.test.2 <- male.stat.test.2 %>% add_xy_position(x = "day", group="dilution")

# show statistics 
male.stat.test
male.stat.test.2

# plot the boxplot with all pairwise stats visualized on top
legend_title <- "Virus Delivery"
ggplot(Male_delta_delta, aes(y= delta_delta_Cq_2, x=day)) + geom_boxplot(outlier.shape=NA,aes(fill=dilution)) +  
  scale_fill_manual(legend_title, values=c("#E7E1EF", "#C994C7", "#CE1256")) + 
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8),aes(fill=dilution)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=10), legend.title=element_text(size=11)) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  labs(title = "Comparing DiNV Increase in Male Flies \nInjected with Various Titers Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection") +
  theme(legend.position = "none")+
  stat_pvalue_manual(male.stat.test, y.position = 7, hide.ns = TRUE)+
  stat_pvalue_manual(male.stat.test.2, y.position = seq(8,12, by=4/4), hide.ns = TRUE)


```

Look at male and female for each dilution by day

3 FFU dilution

```{r}
# male
M_3_dil <- Male_delta_delta[which(Male_delta_delta$dilution == "3 FFU"),]
# female
F_3_dil <- Female_delta_delta[which(Female_delta_delta$dilution == "3 FFU"),]

FFU_3_dil <- rbind(M_3_dil, F_3_dil)

#statistics 
# only care about comparisons between sexes within days 
#doing pairwise t-tests between sexes, within each day
three.stat.test <- FFU_3_dil %>%
  group_by(day) %>%
  t_test(delta_delta_Cq_2 ~ sex) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
#add the bracket location info for each comparison
three.stat.test <- three.stat.test %>% add_xy_position(x = "day")
# show stats 
three.stat.test
# all non significant, so cannot add to the plot 

legend_title = "Sex"
#Plot 
ggplot(FFU_3_dil, aes(y= delta_delta_Cq_2, x=day, fill=sex)) + 
  geom_boxplot()  + 
  scale_fill_manual(legend_title, values=c("#E7298A" ,  "#67001F")) +
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=11), legend.title=element_text(size=12)) +
  theme(legend.position = c(0.9, 0.2), legend.background = element_rect(linetype="solid", colour ="black")) +
  labs(title = "Comparing DiNV Increase in Male and Female Flies \nInjected with 3 FFU DiNV Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection")

```

0.1 FFU dilution

```{r}
# male
M_01_dil <- Male_delta_delta[which(Male_delta_delta$dilution == "0.1 FFU"),]
# female
F_01_dil <- Female_delta_delta[which(Female_delta_delta$dilution == "0.1 FFU"),]

FFU_01_dil <- rbind(M_01_dil, F_01_dil)

#statistics 
# only care about comparisons between sexes within days 
# doing pairwise t-tests between sexes, within each day
o1.stat.test <- FFU_01_dil %>%
  group_by(day) %>%
  t_test(delta_delta_Cq_2 ~ sex) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
# add the bracket location info for each comparison
o1.stat.test <- o1.stat.test %>% add_xy_position(x = "day")
# show stats 
o1.stat.test
# all non significant, so cannot add to the plot 


legend_title = "Sex"
#Plot 
ggplot(FFU_01_dil, aes(y= delta_delta_Cq_2, x=day, fill=sex)) + 
  geom_boxplot()  + 
  scale_fill_manual(legend_title, values=c("#E7298A" ,  "#67001F")) +
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=11), legend.title=element_text(size=12)) +
  theme(legend.position = "none") +
  labs(title = "Comparing DiNV Increase in Male and Female Flies \nInjected with 0.1 FFU DiNV Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection")


```

0.01 FFU dilution

```{r}
# male
M_001_dil <- Male_delta_delta[which(Male_delta_delta$dilution == "0.01 FFU"),]
# female
F_001_dil <- Female_delta_delta[which(Female_delta_delta$dilution == "0.01 FFU"),]

FFU_001_dil <- rbind(M_001_dil, F_001_dil)

#statistics 
# only care about comparisons between sexes within days 
# doing pairwise t-tests between sexes, within each da
oo1.stat.test <- FFU_001_dil %>%
  group_by(day) %>%
  t_test(delta_delta_Cq_2 ~ sex) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj")
# add the bracket location info for each comparison
oo1.stat.test <- oo1.stat.test %>% add_xy_position(x = "day")
# show stats 
oo1.stat.test
# all non significant so not adding to plot 

legend_title = "Sex"
#Plot 
ggplot(FFU_001_dil, aes(y= delta_delta_Cq_2, x=day, fill=sex)) + 
  geom_boxplot()  + 
  scale_fill_manual(legend_title, values=c("#E7298A" ,  "#67001F")) +
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days")) +
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=11), legend.title=element_text(size=12)) +
  theme(legend.position = "none") +
  labs(title = "Comparing DiNV Increase in Male and Female Flies \nInjected with 0.01 FFU DiNV Over Early Infection",y = "Relative DiNV Genome Increase", x = "Days Post Injection")
```


**Adding in days 6, 7, and 9 for males**
Not sure if I will include this, I don't feel comfortable combining two experiments like this and there were larger differences in the day 0 PIF 3 Cqs than I would have expected

Load in dataset

```{r}
Cq_values_extra <- read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/qPCR_analysis/20240228-viral-load-dilutions/20240313-extra-0.01.csv")
```

Look at raw Cq values

```{r}
ggplot(Cq_values_extra, aes(x= Cq, fill = primer)) + geom_histogram(position = "dodge") + facet_grid(~day) 
```

These look pretty similar to above, I wonder if all the flies aren't getting infected...

Calculating extra delta Cq and Plotting

```{r}
# need to organize by name of the sample I think
Cq_values_extra <- Cq_values_extra[order(Cq_values_extra$sample_ID),]
# this worked to order by the sample number 
# and the TPI value is first in the order and PIF 3 second

# Separate that dataframe, incriminating by 2, every number between 1-48 (number of rows in dataframe)
Cq_values_extra$Cq[seq(1,48,2)] # these are the TPI Cq #s
Cq_values_extra$Cq[seq(2,48,2)] # these are the PIF 3 primer Cq #s 

# make the delta Cq by subtracting the PIF 3 values from the TPI primer values
# and this is saved as a vector in R 
delta_Cqs_extra <- Cq_values_extra$Cq[seq(1,48,2)] - Cq_values_extra$Cq[seq(2,48,2)]
#vector
delta_Cqs_extra

# Make a new dataframe that only has one row per sample by getting rid of the rows with the TPI primer
Cq_values_extra_Delta <- Cq_values_extra[which(Cq_values_extra$primer == "PIF 3"),]

# And then add in the delta Cqs as a new column
Cq_values_extra_Delta$delta_Cq <- delta_Cqs_extra

# add a column with 2^ delta Cq
Cq_values_extra_Delta$delta_Cq_2 <- 2^(delta_Cqs_extra)


# plot and use a log 10 scale for the y axis 
ggplot(Cq_values_extra_Delta, aes(y= delta_Cq_2, x=day)) + geom_boxplot()  + theme_linedraw() + geom_point() + scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x)))

```

Delta delta analysis

```{r}
# this might be a little mess to do 
# separate out the day 0 samples to start with 
E_0 <- Cq_values_extra_Delta[which(Cq_values_extra_Delta$day == "day0"),]

# mean of day 0 delta Cq
mean(E_0$delta_Cq)
# -10.274

# main data sheet without day 0 
E <- Cq_values_extra_Delta[which(Cq_values_extra_Delta$day != "day0"),]

# subtract average day 0 delta Cq from the delta Cq of all others 
delta_delta_extra <- E$delta_Cq - -10.274
delta_delta_extra

# add as column 
E$delta_delta_Cq <- delta_delta_extra

# do 2^ delta delta 
E$delta_delta_Cq_2 <- 2^(delta_delta_extra)


# plot
legend_title <- "Virus Delivery"

ggplot(E, aes(y= delta_delta_Cq_2, x=day)) + geom_boxplot() +  
  theme_light() + geom_point() +
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=14)) +
  scale_x_discrete(labels=c("day6" = "6 days", "day7" = "7 days", "day9" = "9 days")) +
  labs(title = "Comparing Viral Titer in Male Flies \nInjected with Various Titers Over Early Infection",y = "2^delta delta Cq", x = "Days Since Injection")
```

Combine this data with the male data from the other experiment

```{r}
# male data
head(M_001_dil)
# to combine datasets I'll need all the same columns 
# don't need plate column, sex column 
M_001_dil_s <- M_001_dil[,c(1,3:4,6:13)]

# combine extra and original data 
# add all of the days back together 
All_Male_delta_delta <- rbind(M_001_dil_s, E)

# plot
legend_title <- "Virus Delivery"

ggplot(All_Male_delta_delta, aes(y= delta_delta_Cq_2, x=day, fill=dilution)) + geom_boxplot() +  
  scale_fill_manual(legend_title, values=c( "#67001F")) + 
  theme_light() + 
  geom_dotplot(binaxis='y', stackdir='center', dotsize=0.75, position=position_dodge(0.8)) + 
  scale_y_continuous(trans='log10', breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) + 
  theme(axis.text=element_text(size=12),axis.title=element_text(size=14), legend.text=element_text(size=12), legend.title=element_text(size=14)) +
  scale_x_discrete(labels=c("day1" = "1 day", "day3" = "3 days", "day5" = "5 days", "day6" = "6 days", "day7" = "7 days", "day9" = "9 days")) +
  labs(title = "Comparing Viral Titer in Male Flies \nInjected with Various Titers Over Infection",y = "2^delta delta Cq", x = "Days Since Injection")
```
