---
title: "P4-Dilutions-combo"
format: gfm
editor: visual
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=300,fig.width=7)
```

Load in packages needed for the analysis

```{r}
library("survival")
library("survminer")
library(dplyr)
library(tidyr)
library(stringr)
```

### Loop to convert the example data.frame 'df' into properly formatted data.frame 'results'

```{r}
#write a function to transform a data.frame that has the column format 'vial | treatment | D0 | D1 | D2...', with one row for each vial
#into a long version in tidy format that can be input to make a survivorship curve
convert_df<-function(df){
  #open empty data.frame to store results
  results<-data.frame(vial=character(),treatment=character(),dead=numeric(),status=numeric())
  #This loop will index out each row (one row per vial) one at a time, transform it into long format (one row per fly), and add the information to the empty data.frame called results
  for(i in 1:nrow(df)){
  #isolate the row (vial) you want to work on
  temp<-df[i,]
    #iteratively isolate each day for this vial (day 0 must be column 3, day 1 column 4, etc.). Loop stops the column before the last day
    for(j in 3:(ncol(temp)-1)){
      #assign the number of flies that died in the vial on that day (starting with day 1) to the variable 'z'
      z<-temp[1,j]-temp[1,j+1]
        #if >0 flies died add this information to the results dataframe
        if(z>0){
          #iterate over 1 through total number of dead flies
          for(k in 1:z){
            #add a new row to the 'results' data.frame for the given dead fly, specifying vial #, treatment, day died, and
            #record the current vial #
            vial<-temp[,1]
            #record the genotype of the current vial
            treatment<-temp[,2]
            #record the death date of the flies that died on this day (assumes that your input DF starts with day 0 in column 3)
            dd<-j-2
            #append this information into a new row in the 'results' data.frame, and add a '1' in the 4th column to indicate mortality
            results[nrow(results)+1,]<- c(vial,treatment,dd,1)
          } #close for loop
        } #close if loop
    } #close for loop
  
  #now assign the number of flies remaining in the vial on the last day (value in the last column of the row) to the variable 'z'
  z<-temp[1,j+1]
    #if there are any flies alive in the vial on the last day
    if(z>0){
      #iterate over 1:(number of flies alive on the last day)
      for(l in 1:z){
        #record the current vial #
        vial<-temp[,1]
        #record the genotype of the current vial
        treatment<-temp[,2]
        #record the last day we recorded this fly alive (assumes that your input DF starts with day 0 in column 3)
        dd<-j-2
        #append this information into a new row in the 'results' data.frame, and add a '0' in the 4th column to indicate that the fly made it to the end of the experiment
        results[nrow(results)+1,]<- c(vial,treatment,dd,0)
      } #close for loop
    } #close if loop
  } #close original for loop
results$dead<-as.numeric(results$dead)  #reiterate that this column must be class numeric
results$status<-as.numeric(results$status)  #reiterate that this column must be class numeric
results$vial <- as.factor(results$vial) # make sure vial is considered a factor
# gives you only the results dataframe as output from function 
return(results) 
} #close function
```

Read in raw data

**Note that for these datasets, the adjusted N number is the number of flies alive on day 2 because there seemed to be more early death in these experiments than usual**

```{r}
#read the file from csv
df<-read.csv("/Users/maggieschedl/Desktop/Github/Unckless_Lab_Resources/Infection_survival_analyses/20231205-dilution-rep-3/dilution-infection-combo-sheet.csv")

# Remove the rows of the undiluted samples because I did not do those for the other replicates
df <- df[c(1:12, 15:42),]
# separate out columns needed
df <- df[,c(1,3,15:29)]

```

Convert dataframe

```{r}
df.convert<-convert_df(df)
```

Plot survivial curve with every line separate, yes this is a mess

```{r}
# change to not have confidence intervals in this one so you can see them 
df_fit<- survfit(Surv(dead, status) ~ treatment, data=df.convert)
ggsurvplot(df_fit,
          pval = FALSE, conf.int = FALSE,
          #risk.table = TRUE, # Add risk table
          #risk.table.col = "strata", # Change risk table color by groups
          #linetype = "strata", # Change line type by groups
          #surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw()) # Change ggplot2 theme
          # palette = c("orchid", "aquamarine", "blueviolet", "darkslategray3")) + ylab("Survival Proporation") + xlab("Days post injection")

```

Add in columns for sex and dilution separately

```{r}
# add in sex information and DiNV information as separate columns by splitting the treatment column 
# split the columns 
df.convert_S <- str_split_fixed(df.convert$treatment, "-", 2)

# change column names
colnames(df.convert_S) <- c("sex", "dilution")

# add columns to df 
df.convert_full <- cbind(df.convert,df.convert_S)

# add in block information 
# this was not reproducable to count but I don't know how else to do it
# first 118 rows are block A, second 127 rows are block B, and last 138 rows are block C
df.convert_full$Block <- rep(c("A","B", "C"), c(118, 127, 138))

```

Plot survival curve by dilution with the sexes combined

```{r}
# change to not have confidence intervals in this one so you can see them 
df2_fit<- survfit(Surv(dead, status) ~ dilution, data=df.convert_full)
ggsurvplot(df2_fit,
          pval = FALSE, conf.int = FALSE,
          #risk.table = TRUE, # Add risk table
          #risk.table.col = "strata", # Change risk table color by groups
          #linetype = "strata", # Change line type by groups
          #surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw()) # Change ggplot2 theme
          # palette = c("orchid", "aquamarine", "blueviolet", "darkslategray3")) + ylab("Survival Proporation") + xlab("Days post injection")

```

Plot survival curve with only females

```{r}

df.convert_fem <- df.convert_full[which(df.convert_full$sex != "male"),]

# change to not have confidence intervals in this one so you can see them 
df3_fit<- survfit(Surv(dead, status) ~ dilution, data=df.convert_fem)
ggsurvplot(df3_fit,
          pval = FALSE, conf.int = FALSE,
          #risk.table = TRUE, # Add risk table
          #risk.table.col = "strata", # Change risk table color by groups
          #linetype = "strata", # Change line type by groups
          #surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw()) # Change ggplot2 theme
          # palette = c("orchid", "aquamarine", "blueviolet", "darkslategray3")) + ylab("Survival Proporation") + xlab("Days post injection")

```

Plot survival curve with only males

```{r}

df.convert_m <- df.convert_full[which(df.convert_full$sex != "female"),]

# change to not have confidence intervals in this one so you can see them 
df4_fit<- survfit(Surv(dead, status) ~ dilution, data=df.convert_m)
ggsurvplot(df4_fit,
          pval = FALSE, conf.int = FALSE,
          #risk.table = TRUE, # Add risk table
          #risk.table.col = "strata", # Change risk table color by groups
          #linetype = "strata", # Change line type by groups
          #surv.median.line = "hv", # Specify median survival
          ggtheme = theme_bw()) # Change ggplot2 theme
          # palette = c("orchid", "aquamarine", "blueviolet", "darkslategray3")) + ylab("Survival Proporation") + xlab("Days post injection")

```

Models

```{r}
# standard model with both sexes
df_fit_combo<- coxph(Surv(dead, status) ~ treatment, data=df.convert_full)
summary(df_fit_combo)
```

```{r}
# standard model only males
df_fit_combo_m<- coxph(Surv(dead, status) ~ treatment, data=df.convert_m)
summary(df_fit_combo_m)
```

```{r}
# standard model only females
df_fit_combo_f<- coxph(Surv(dead, status) ~ treatment, data=df.convert_fem)
summary(df_fit_combo_f)
```

```{r}
# model with block both sexes
df_fit_combo_b<- coxph(Surv(dead, status) ~ treatment + Block, data=df.convert_full)
summary(df_fit_combo_b)
```

```{r}
# model with block males
df_fit_combo_b_m<- coxph(Surv(dead, status) ~ treatment + Block, data=df.convert_m)
summary(df_fit_combo_b_m)
```

```{r}
# model with block females
df_fit_combo_b_f<- coxph(Surv(dead, status) ~ treatment + Block, data=df.convert_fem)
summary(df_fit_combo_b_f)
```


Changing how to name the dilutions 
```{r}
df.convert_full$dilution2 <- df.convert_full$dilution
df.convert_full$dilution2[which(df.convert_full$dilution2 == "CCM")] = 0
df.convert_full$dilution2 = as.numeric(df.convert_full$dilution2)

hist(df.convert_full$dilution2)

```

```{r}
# model considering sex by dilution 
df_fit_combo_robs_model <- coxph(Surv(dead, status) ~ sex * dilution2 + Block, data=df.convert_full)
summary(df_fit_combo_robs_model)
```

Run model without control flies 

```{r}
df.convert_full2 <- df.convert_full[which(df.convert_full$dilution2 != 0),]
# model considering sex by dilution 
df_fit_combo_robs_model_2 <- coxph(Surv(dead, status) ~ sex * dilution2 + Block, data=df.convert_full2)
summary(df_fit_combo_robs_model_2)
```
